Thesis Title

### 2022-05-23 tables/chapter_2/table3.tex
mv tables in chapter 2 to folder

\caption{Microorganisms identified by conventional methods, WGS and using shotgun metagenomics and the taxonomic classification methods in Unix.}
\label{tab:ch2_table2}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}|l|l|l|l|lll|@{}}
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{WGS-based identification}}} &
  \multicolumn{3}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){5-7} 
  \multicolumn{1}{c|}{\textbf{Kraken$^b$}} &
  \multicolumn{1}{c|}{\textbf{MIDAS$^c$}} &
  \multicolumn{1}{c|}{\textbf{MetaPhlAn$^c$}} \\ \midrule
  \begin{tabular}[c]{@{}l@{}}10$^3$\\  10$^3$\\  10\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (34.6\%)\\  \textit{S. haemolyticus} (10.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (62.0\%)\\  \textit{S. haemolyticus} (28.0\%)\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (66.6\%)\\  \textit{S. haemolyticys} (27.7\%)\\  -\end{tabular} \\ \midrule
  \begin{tabular}[c]{@{}l@{}}-\#\\  -\#\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (29.5\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (100.0\%)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (100.0\%)\end{tabular} \\ \midrule
  \multicolumn{1}{l|}{S. aureus (0.2\%)} &
  Not identified$^*$ \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (0.73\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus }(100\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  -$^\#$\\  \textit{E. faecalis}\\  -$^\#$\\  -$^\#$\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (9.7\%)\\  \textit{K. oxytoca} (0.5\%)\\  \textit{S. anginosus} (0.07\%)\\  \textit{E. faecalis} (0.3\%)\\  Several species (12.7\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (6.5\%)\\  \textit{K. oxytoca} (0.3\%)\\  \textit{S. anginosus} (0.01\%)\\  \textit{E. faecalis} (0.9\%)\\  Several species (96.7\%)\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli} (8.5\%)\\  \textit{K. oxytoca} (0.3\%)\\  \textit{Streptococcus spp.} (0.09\%)\\  \textit{E. faecalis} (0.7\%)\\  Several species (90.4\%)\\  -\end{tabular} \\ \midrule
  \multicolumn{1}{l|}{\textit{E. faecium} (0.77\%)} &
  Not identified$^*$ \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (82.9\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{O. anthropi} (21.3\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (99.4\%)} &
  \textit{O. intermedium} (99.1\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (22.9\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{S. marcescens} (64.7\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (99.1\%)} &
  \textit{S. marcescens} (100\%) \\ \bottomrule
\tiny
\item $^d$miniKraken database was used
\item $^\#$Although there was a laboratory identification, no isolates were available for WGS
\caption{Microorganisms identified by conventional methods, WGS and using shotgun metagenomics and the taxonomic classification methods in CLC Genomics Workbench.}

@@

\caption{Microorganisms identified by conventional methods, WGS and using shotgun metagenomics and the taxonomic classification methods in webpages (BaseSpace, Taxonomer and CosmosID).}
\label{tab:ch2_table4}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}|l|l|l|l|lllll|@{}}
\multirow{2}{*}{\textbf{Sample number}} &
  \multirow{2}{*}{\textbf{Culture result (CFU)$^a$}} &
  \multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}} &
  \multirow{2}{*}{\textbf{WGS-based identification}} &
  \multicolumn{5}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){5-9} 
 &
   &
   &
   &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Genius\\  (Basespace)$^c$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}Kraken\\  (Basespace)$^c,d$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{\begin{tabular}[c]{@{}l@{}}MetaPhlAn\\  (Basespace)$^c$\end{tabular}}} &
  \multicolumn{1}{l|}{\textbf{Taxonomer (Utah)$^b,e$}} &
  \textbf{Cosmos ID$^a$} \\ \midrule
  \begin{tabular}[c]{@{}l@{}}10$^3$\\  10$^3$\\  10\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (14.4\%)\\  \textit{S. haemolyticus} (55.8\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (25.0\%)\\  \textit{S. haemolyticus} (20.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (65.1\%)\\  \textit{S. haemolyticys} (30.4\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (22.9\%)\\  \textit{S. haemolyticus} (20.1\%)\\  Not identified$^*$\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium} (50.3\%)\\  \textit{S. haemolyticus} (22.1\%)\\  \textit{C. glabrata} (88.6\%)\end{tabular} \\ \midrule
  \begin{tabular}[c]{@{}l@{}}103\\  1\\  Not determined\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (94.0\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (27.0\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (54.2\%)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (14.2\%)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not identified$^*$\\  Not identified$^*$\\  Several species (100\%)\end{tabular} \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (0.1\%)} &
  \multicolumn{1}{l|}{\textit{S. pseudintermedius} (3.4\%)} &
  Not identified$^*$ \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (0.3\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (8.3\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  -$^\#$\\  \textit{E. faecalis}\\  -$^\#$\\  -$^\#$\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (0.4\%)\\  Not identified$^*$\\  \textit{S. anginosus} (0.03\%)\\  \textit{E. faecalis} (0.8\%)\\  Several species (45.0\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (10.2\%)\\  \textit{K. oxytoca} (0.5\%)\\  \textit{S. anginosus} (0.4\%)\\  \textit{E. faecalis} (0.3\%)\\  Several species (8.0\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (7.0\%)\\  \textit{K. pneumoniae} (0.01\%)\\  \textit{S. anginosus} (0.3\%)\\  \textit{E. faecalis} (0.7\%)\\  Several species (89.1\%)\\  -\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}\textit{E. coli} (3.6\%)\\  \textit{K. michiganensis} (0.1\%)\\  \textit{S. anginosus} (0.1\%)\\  \textit{E. faecalis} (0.1\%)\\  Several species (60.3\%)\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli} (7.6\%)\\  \textit{K. oxytoca} (1.7\%)\\  \textit{S. anginosus} (0.09\%)\\  \textit{E. faecalis} (3.7\%)\\  Several species (86.2\%)\\  Not identified$^*$\end{tabular} \\ \midrule
  \multicolumn{1}{l|}{\textit{E. faecium} (4.2\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (14.8\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (5.5\%)} &
  \multicolumn{1}{l|}{\textit{E. faecium} (1.4\%)} &
  \textit{E. faecium} (4.1\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (93.8\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (14.2\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{O. intermedium} (100\%)} &
  \multicolumn{1}{l|}{\textit{O. nthropic} (88.9\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (99.8\%)} &
  \multicolumn{1}{l|}{\textit{O. intermedium} (13.1\%)} &
  \textit{O. intermedium} (49.5\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{S. aureus} (99.5\%)} &
  \multicolumn{1}{l|}{\textit{S. aureus} (12.7\%)} &
  \textit{S. aureus} (100\%) \\ \midrule
  \multicolumn{1}{l|}{\textit{S. marcescens} (32.5\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (94.8\%)} &
  \multicolumn{1}{l|}{\textit{Serratia} spp. (100\%)} &
  \multicolumn{1}{l|}{\textit{S. marcescens} (1.4\%)} &
  S\textit{. marscescens} (38.4\%) \\ \bottomrule
\tiny
\item $^d$miniKraken database was used
\itel $^e$Full Analysis mode was used 
\item $^\#$Although there was a laboratory identification, no isolates were available for WGS

\subsection{Comparison of standard procedures and shotgun metagenomics for the identification of clinically relevant pathogens}

The tools used for taxonomic classification are shown in Figure \ref{chap2_figure1}. 
The total number of reads and the total number of reads mapped against the human genome (hg19) varied between samples, ranging from 3.5\% to 98.9\% (Table \ref{tab:ch2_table_1}). 
The abundance of human reads was not determined by the type of sample but was probably influenced by individual characteristics of each sample and the success of the methods used in depleting the human DNA. 
We identified the microorganisms present using different taxonomical methods, including three Unix-based tools (Kraken, Metaphlan2 and MIDAS), web-based tools including both commercial and freely available solutions (BaseSpace, Taxonomer and CosmosID) and one commercial approach having a graphical interface (CLC Genomics Workbench v10.0.1). 
The taxonomic classification results for each sample are presented in Tables \ref{tab:ch2_table2}, \ref{tab:ch2_table3} and \ref{tab:ch2_table4}. 
In 8 samples, all the microorganisms identified by classical culture were also identified through metagenomics (using at least one method). 
In sample 2, two of the bacterial species identified by classical culture, i.e., \textit{E. coli} and one \textit{Enterococcus avium} were not identified through shotgun metagenomics and in sample 3 there was no concordance between the results of MALDI-TOF and the taxonomical classification methods at the species level (Tables \ref{tab:ch2_table2}, \ref{tab:ch2_table3} and \ref{tab:ch2_table4}).
We identified \textit{Ochrobactrum intermedium} in the negative control, but in low amounts (1.0\% of the reads mapped to the reference genome with the accession number NZ\_ACQA01000002 and only 1.4\% of the reference genome was covered). 
The sensitivity and positive predictive value of each classification method is shown in Table \ref{ch2_table5}.

\begin{table}[]
\caption{Performance of the different taxonomic classification methods for each sample. Sensitivity and positive predictive value were calculated using culture/MALDI-TOF as standards}
\label{tab:ch2_table5}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lllllll@{}}
\toprule
\textbf{Method} &
  \textbf{Total number of bacteria identified$^a$} &
  \textbf{True positives$^a$} &
  \textbf{False positives} &
  \textbf{False negatives} &
  \textbf{Sensitivity (\%)} &
  \textbf{PPV (\%)} \\ \midrule
Culture/MALDI-TOF                                         & 9    & 9 & 0    & 0 & 1    & 1    \\
MetaPhlAn (BaseSpace)                                     & 16   & 7 & 9    & 2 & 0.78 & 0.44 \\
Genius (BaseSpace)                                        & 35   & 8 & 27   & 1 & 0.89 & 0.23 \\
Kraken (BaseSpace)                                        & 959  & 7 & 952  & 2 & 0.78 & 0.01 \\
Taxonomer (Full Analysis)                                 & 4649 & 8 & 4641 & 1 & 0.89 & 0    \\
CosmosID                                                  & 35   & 8 & 27   & 1 & 0.89 & 0.23 \\
Taxonomic Profiling (CLC Genomics Workbench v10.0.1)      & 17   & 6 & 11   & 3 & 0.67 & 0.35 \\
Best match K-mer spectra (CLC Genomics Workbench v10.0.1) & 12   & 8 & 4    & 1 & 0.89 & 0.67 \\
Kraken (Unix)                                             & 198  & 7 & 191  & 2 & 0.78 & 0.04 \\
MetaPhlAn2 (Unix)                                         & 15   & 7 & 6    & 4 & 0.75 & 0.75 \\
MIDAS (Unix)                                              & 34   & 7 & 26   & 2 & 0.88 & 0.5 \\ \bottomrule
\end{tabular}%
}
\tiny
\item $^a$aExcluding the samples with non-identified anaerobic bacteria (Samples 2 and 5)
\item Abbreviations: PPV – positive predictive value
\end{table}

\begin{figure*}[h!]
\centering
\includegraphics[angle=-90,width=\textwidth]{figures/chapter 2/41598_2018_31873_Fig1_HTML.pdf}
\caption{Scheme of the bioinformatic analysis of the metagenomics samples.}
\label{fig:chap2_figure1}
\end{figure*}

\subsection{Determination of antimicrobial resistance}

Metagenomics provides other sequence information in addition to pathogen detection. 
We determined the presence of antimicrobial-resistance genes in the SMg sequence data and compared the results with those obtained from WGS and phenotypic resistance testing (Table \ref{ch2_table6}).

AMR genes found with CLC Genomics Workbench and ReMatCh in samples 1, 7 and 9 correlated well with phenotypic results. 
However, in the other 7 samples, not all antimicrobial resistance genes that could explain the phenotypic profile were identified. 
In addition, in samples 2, 5, 7 and 10, ReMatCh detected different resistance genes compared to those reported by CLC Genomics Workbench (Table \ref{ch2_table6}).
Some of these differences (genes \textit{norA}, \textit{blaSST}-1, \textit{fusA}) were due to slight differences in the databases used, however, the other resistance genes were present in both databases. 
Interestingly, in two samples (samples 2 and 5), we were able to identify several antimicrobial resistance genes usually found in anaerobic bacteria. 
These were not reported by classical microbiology methods, probably because they were not considered relevant pathogens worthy of subsequent susceptibility study (mixed anaerobic culture).

The SEAR app in BaseSpace (the only one available for antimicrobial resistance gene detection) crashed several times, although we performed the analysis repeatedly, using different parameters.
We were only able to get results in 3 samples, with no resistance genes detected.
\caption{Antimicrobial resistance phenotypes and antimicrobial resistance genes detected using different approaches.}
\label{tab:ch2_table6}
\resizebox{\textwidth}{!}{%
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional susceptibility testing\\  (VITEK 2)b\end{tabular}}}} &
  \multicolumn{1}{c|}{\textbf{WGS}} &
  \multicolumn{2}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){4-6} 
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbench}} &
  \multicolumn{1}{c|}{\textbf{ReMatCh (Unix)}} &
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbencha}} \\ \midrule
  \begin{tabular}[c]{@{}l@{}}E. faecium\\  S. haemolyticus\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}LEV, ERY, CLI\\  OXA, GEN, CIP, FOS, ERY, CLI\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, dfrG\\  blaZ, mecA, ant(6’)-Ia, aph(3’)-III, aac(6’)-aph(2’’), erm(C), mph(C), msr(A), dfrG\end{tabular} &
  \multicolumn{1}{l|}{erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, aac(6’)-aph(2’’), blaZ, mecA, erm(C), mph(C), msr(A), dfrG} &
  erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, aac(6’)-aph(2’’), blaZ, mecA, erm(C), mph(C), msr(A), dfrG \\ \midrule
  \begin{tabular}[c]{@{}l@{}}E. avium\\  E. coli\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}DOX, CLI\\  susceptible\\  -\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-\#\\  -\#\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not detected\\  Not detected\\  catS, lnu(D), lsa(C), cepA-44, tet(Q)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}Not detected\\  Not detected\\  catS, lnu(D), lsa(C), cepA-44, tet(Q), fusA\end{tabular} \\ \midrule
  S. epidermidis &
  OXA, GEN, TEC, FUS, CIP, ERY, CLI &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \midrule
  S. aureus &
  PEN, ERY &
  blaZ, spc, erm(A) &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \midrule
  \begin{tabular}[c]{@{}l@{}}E. coli\\  K. oxytoca\\  S. anginosus\\  E. faecalis\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}susceptible\\  AMX\\  susceptible\\  DOX, CLI\\  -\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-\#\\  blaOXY-1-3\\  -\#\\  tet(M), lsa(A)\\  -\#\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\\  tet(M)\\  cfxA4, tet(Q)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\\  tet(O)\\  cfxA4, tet(Q)\end{tabular} \\ \midrule
  E. faecium &
  PEN, AMX, CFX, IMP, GENhl, STRhl, LEV, ERY, CLI, AMP/SUL &
  erm(B), msr(C), ant(6’)-Ia, aph(3’)-III, aac(6’)-aph(2’’), dfrG &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \midrule
  S. aureus &
  PEN &
  blaZ &
  \multicolumn{1}{l|}{blaZ, norA} &
  blaZ \\ \midrule
  O. intermedium &
  AMX, PIP/TAZ, CFX, CFT, CTZ, IMP, FOX, TOB, FOS, NIT, TMP &
  blaOCH-2 &
  \multicolumn{1}{l|}{blaOCH-5} &
  blaOCH-2 \\ \midrule
  S. aureus &
  PEN &
  \multicolumn{1}{l|}{blaZ} &
  blaZ \\ \midrule
  S. marcescens &
  AMX, AMC, CFX, FOX, NIT, POL &
  \multicolumn{1}{l|}{blaSST-1, tet(41), oqxB, aac(6’)-Ic} &
  tet(41), oqxB, aac(6’)-Ic \\ \bottomrule
\item $^a$The analysis aborted when the script tried to connect to NCBI
\item $^b$Only non-susceptibility is indicated. \item Abbreviations: AMP/SUL, ampicillin/sulbactam; AMX, amoxicillin; AMC, amoxicillin/clavulanate; CFX, cefuroxime; FOS, fosfomycin; FOX, cefoxitin; CIP, ciprofloxacin; CLI, clindamycin; DOX, doxycycline; ERY, erythromycin; FUS, fusidic acid; GEN, gentamicin; GENhl, gentamicin high-level; LEV, levofloxacin; NIT, nitrofurantoin; PEN, penicillin; POL, polymyxin B; STRhl, streptomycin high-level; TEC, teicoplanin.

\subsection{MLST and wgMLST analysis}

In three cases when SMg data covered $\geq$ 93\% of the genome we were able to identify the ST, which corresponded to the one found using WGS of the isolated bacteria using CLC Genomics Workbench (n=2) and metaMLST (n=1). 
These results are summarized in Table \ref{tab:ch2_table7}.
Assembled genomes and metagenomes, were compared by wgMLST analysis using Ridom SeqSphere+. 
Figure \ref{fig:chap2_figure2} shows examples of the allele difference between the genomes obtained through WGS versus the genomes obtained through shotgun metagenomics.

\begin{table}[]
\caption{Results of MLST using by whole genome sequencing and shotgun metagenomics}
\label{tab:ch2_table7}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}|l|l|l|ll|@{}}
\toprule
\multicolumn{1}{|c|}{\multirow{2}{*}{\textbf{Sample number}}} &
  \multicolumn{1}{c|}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}Conventional identification\\  (MALDI-TOF)\end{tabular}}}} &
  \multicolumn{1}{c|}{\textbf{WGS}} &
  \multicolumn{2}{c|}{\textbf{Shotgun metagenomics}} \\ \cmidrule(l){3-5} 
\multicolumn{1}{|c|}{} &
  \multicolumn{1}{c|}{} &
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbench v10.1.1}} &
  \multicolumn{1}{c|}{\textbf{CLC Genomics Workbench v10.1.1}} &
  \multicolumn{1}{c|}{\textbf{metaMLST (Unix-based)}} \\ \midrule
\textbf{1} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. faecium}\\  \textit{S. haemolyticus}\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}ST117\\  ST25\end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Not detected (6 alleles identified correctly)\\  Not detected (3 alleles identified correctly)\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}ST117\\  Not detected\end{tabular} \\ \midrule
\textbf{2} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. avium}\\  \textit{E. coli}\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}-$^\#$ \\  -$^\#$ \\  -$^\#$ \end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}-\\  Not detected\\  -\end{tabular} \\ \midrule
\textbf{3} &
  \textit{S. epidermidis} &
  -$^\#$ &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \midrule
\textbf{4} &
  \textit{S. aureus} &
  ST30 &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \midrule
\textbf{5} &
  \begin{tabular}[c]{@{}l@{}}\textit{E. coli}\\  \textit{K. oxytoca}\\  \textit{S. anginosus}\\  \textit{E. faecalis}\\  Anaerobes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}ST141\\  ST40\\  -$^\#$ \\  ST179\\  -$^\#$ \end{tabular} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}ST141\\  Not detected\\  -\\  Not detected\\  -\end{tabular}} &
  \begin{tabular}[c]{@{}l@{}}ST4508\\  Not detected\\  -\\  Not detected\\  -$^\#$ \end{tabular} \\ \midrule
\textbf{6} &
 \textit{ E. faecium} &
  ST117 &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \midrule
\textbf{7} &
  \textit{S. aureus} &
  ST30 &
  \multicolumn{1}{l|}{ST30} &
  ST667 \\ \midrule
\textbf{8} &
  \textit{O. intermedium} &
  - &
  \multicolumn{1}{l|}{-} &
  - \\ \midrule
\textbf{9} &
  \textit{S. aureus} &
  -$^\#$  &
  \multicolumn{1}{l|}{Not detected} &
  Not detected \\ \midrule
\textbf{10} &
  \textit{S. marcescens} &
  -$^\#$ &
  \multicolumn{1}{l|}{-} &
  - \\ \bottomrule
\end{tabular}%
}
\item Abbreviations: ST, sequence type
\end{table}

\begin{figure*}[h!]
\centering
\includegraphics[angle=-90,width=\textwidth]{figures/chapter 2/41598_2018_31873_Fig2_HTML.pdf}
\caption{Minimum-spanning tree based on wgMLST allelic profiles of 2 S. aureus genomes and 2 E. coli genomes obtained through SM and WGS in comparison to reference strains 04-02981 (GenBank accession number NC\_017340) and 06-00048 (NZ\_CP015229), respectively. Each circle represents an allelic profile based on sequence analysis. The numbers on the connecting lines illustrate the numbers of target genes with differing alleles.}
\label{fig:chap2_figure2}
\end{figure*}

\subsection{Characterisation of mobile genetic elements}

Two different approaches, i.e. CLC Genomics Workbench and Bowtie2 were used to identify plasmids present in the sequence data. 
Both approaches used mapping of sequences against the same plasmid database. 
Since some plasmids present in the database are very similar and sequence reads may be mapped to more than one plasmid, we used the pATLAS tool, which provides an overview of the nodes (representing plasmid sequences) and links between plasmids (which connect similar plasmids), to enable the visualisation of the plasmids identified (Figure \ref{fig:chap2_figure3}). 
A colour gradient indicates the sequence coverage of the plasmids. 
In most cases, the same plasmids were identified by both approaches, with some small differences in sequence coverage. 
When comparing the plasmids identified in the SMg dataset versus the WGS data, most of the plasmids were also detected in the isolates (an example is shown in Figure \ref{fig:chap2_figure4}). 
However, some plasmids were not identified in any of the isolated bacteria and were probably residing in low-abundant species.

\begin{figure*}[h!]
\centering
\includegraphics[angle=-90,width=\textwidth]{figures/chapter 2/41598_2018_31873_Fig3_HTML.pdf}
\caption{(a) Overview of the nodes (representing plasmid sequences) and links between plasmids (connecting similar plasmids) found in Sample 1 (SMg) using the pATLAS tool. (b) A closer look at one of the cloud of plasmids. The colour gradient in each cloud of plasmids represents the plasmid sequence coverage (SC), varying between 0-0.79 (grey) and 0.80-1 (red gradient).}
\label{fig:chap2_figure3}
\end{figure*}

\begin{figure*}[h!]
\centering
\includegraphics[angle=-90,width=\textwidth]{figures/chapter 2/41598_2018_31873_Fig4_HTML.pdf}
\caption{A heatmap comparing the identified plasmids using bowtie2 in \textit{S. haemolyticus} WGS (1),\textit{ E. faecium} WGS (2) and in the SMg dataset (3) isolated from sample 1.}
\label{fig:chap2_figure4}
\end{figure*}

\section{Discussion}

This study evaluated the suitability of SMg for the microbiological diagnosis and (patho- and epi-) typing of microorganisms directly from real patient samples. The whole procedure took between 48-54 hours to complete, which is shorter than culture-based methods if one includes typing. However, the amount of information derived from SMg in most cases, did not overcome the necessity for pathogen isolation and subsequent (phenotypic and genotypic) typing, which can take up to 1-2 weeks (particularly in slow-growing organisms). Nevertheless, SMg can help guide antimicrobial therapy and be helpful in cases where there is a suspicion of transmission and there is a need to quickly determine the genetic relationship between pathogens, although the success of SMg in individual patient samples can be highly variable, as reported here.

Different bioinformatics pipelines were evaluated to identify potential differences between them and identify those which could provide the clinical microbiologist with the maximum of relevant and accurate information. In terms of microbial identification, in both Unix and web-based approaches we would recommend MetaPhlAn, since it has good sensitivity and a good positive predictive value (PPV). The find best match K-mer spectra tool should be used in the context of the CLC Genomics Workbench, since it had a higher sensitivity and PPV compared to the Taxonomic Profiling tool.

In a clinical setting, a combination of high sensitivity and high PPV of any new method is key. Popular software designed for bacterial identification, can predict dozens to hundreds of species in in vitro generated bacterial communities of known composition \citep{peabody_evaluation_2015}. We observed the same when using Kraken and Taxonomer when comparing to culture-based methods. For both Kraken and Taxonomer, relative abundance cut-off values may be required to limit the number of species identified. However, which cut-off values should be used are a matter of debate, since in some cases, even if applying a cut-off value as low as 1.0\% (comparable to what was found in the negative control) would have resulted in decreased sensitivity (e.g. the \textit{Streptococcus anginosus} identified by culture in Sample 5 would have been disregarded). The methods that employ several parameters to infer microbial identification are superior, because they not only rely on the relative abundance of bacterial species, but also on the genome coverage and on the proportion of the genome that was covered. On the other hand, in some cases SMg may be more sensitive than culture in identifying pathogens, reflecting the higher sensitivity or the capacity to detect bacterial species which are non-culturable in the conditions used or that are no longer culturable, such as due to prior antimicrobial therapy. In such cases, other methods like 16S rDNA sequencing or the recently described 16S-23S rDNA sequencing method \citep{sabat_targeted_2017} may be used for discrepancy analyses. However, here we decided to use culture-based methods as the gold standard, since this is still the method of choice in clinical microbiology.

One limitation of this study was the exclusion of culture-negative samples and thus their inclusion would have affected the calculation of the specificity values. However, as mentioned above, culture-negative samples do not necessarily mean that the samples are pathogen-free, but it might only reflect the low sensitivity or capacity of culture-based methods to detect non-culturable bacterial species. As with other (molecular) methods, several controls should be included to validate the obtained results, including a negative control. In our negative control, we detected an \textit{O. intermedium} strain, although with only 1.0\% of the reads mapping to the reference genome and covering only 1.4\% of the reference genome (accession number NZ\_ACQA01000002). These results may be due to contamination during library preparation (e.g. sample-to-sample contamination prior to indexing), the result of sequencing artefacts (e.g. demultiplexing errors), or to incorrect classification during data analysis (e.g. highly similar regions) \citep{graf_unbiased_2016}. Our samples and sequencing libraries were handled in laminar flow cabinets; however, we cannot also exclude the possibility of contamination. Furthermore, the reagents used may also be or become contaminated with DNA leading the detection of these contaminating species, something that has been described previously \citep{street_molecular_2017}. This poses a challenge for interpretation, because some positive samples also had very low numbers of reads for some pathogens (< 1\%). When approaching this limit of detection, small numbers of pathogen reads will be difficult to interpret, as they can represent true-positives with low abundance in the sample, or artefacts such as contamination during library preparation\citep{graf_unbiased_2016}.

In terms of antimicrobial resistance gene detection, ReMatCh (Unix) and the CLC Genomics Workbench Find Resistance tool gave comparable results. Since ReMatCh (Unix) performs the analysis at the read level, while CLC Genomics Workbench performs it at the contig level, we suggest that both strategies should be employed in parallel when looking for antimicrobial resistance genes. It is also important to emphasise that the contig-level approach employed by CLC Genomics Workbench may give negative results if the sequence coverage is set to a high percentage (e.g. above 80\%). This is due to the assembly method, which may split the antimicrobial resistance genes into different contigs, when the number of reads is too low. This phenomenon was observed in Sample 1, for the \textit{aac(6’)-aph(2’’)} gene, which was split into 3 different contigs, each part corresponding to less than 40\% of the gene. Only when applying a cut-off value of $geq$ 20\% for sequence coverage could we identify all three parts of the gene, which in total corresponded to 89\% of the entire sequence. Finally, it is important to point out that the ResFinder database (used here), and other databases, focus on acquired genes, not including chromosomal point mutations resulting in antimicrobial resistance. However, a recently developed tool, PointFinder, was added to ResFinder for the detection of chromosomal point mutations associated with antimicrobial resistance \citep{zankari_pointfinder_2017} and an updated database will be available soon.

>>>>

\caption{Microorganisms identified by conventional methods, \ac{WGS} and using shotgun metagenomics and the taxonomic classification methods in CLC Genomics Workbench.}

@@
